### 2019-03-11 2-Chapters/4-Chapter/chapter4.tex
Commence à intégrer changements d'Emilie, cf. #13
Should be good for Chapter 4 :boom:

More precisely, we start by motivating this idea of independent ``selfish'' learning for IoT devices in Section~\ref{sec:4:motivations}.
We then present in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, occupied by static devices.
Devices do not use any sensing, but use acknowledgements sent by the Base Station to know if their uplink messages were successfully received.
We consider the case when many independent ``dynamic'' devices all have a small probability of communicating at every discrete time. If they use simple MAB algorithms in a decentralized and independent fashion (what we call ``selfish'' learning),
we show in simulations that in the case of a non-uniform repartition of the static devices in the channels, the dynamic devices can drastically improve the network efficiency.
%
On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real wireless radio hardware.

We then present in Section~\ref{sec:4:retransmissions} a second model extending our first model, when devices can retransmit their messages for a fixed number of maximum retransmission (like in the ALOHA protocol).
In this second model, we compare five different heuristics, all based on the UCB algorithm, and numerical simulation show that all the non-naive heuristics succeed in significantly improving the network efficiency.
Interestingly, the simpler heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.

@@

This Chapter is mainly based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.

@@

Efficient Medium Access (MAC) policies allow devices to avoid interfering traffic and can significantly reduce the spectrum contention problem in unlicensed bands.
As end-devices battery life is a key constraint of IoT networks,
this leads to IoT protocols using as low signaling overhead as possible and simple ALOHA-based mechanisms.
%
In this Chapter, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, used in combination with a time-frequency slotted ALOHA-based protocol.
We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
Kaufmann12Thompson}, for the first model. For the demonstration as well as for the second model, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored for solve the problem of learning in order to retransmit efficiently.

@@

We present in Section~\ref{sec:4:firstModel} how the MAB algorithms can be used in a unlicensed but frequency- and time-slotted IoT network.
Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously and their activation processes are random, their behavior is not stationary.
As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``highly'' not stationary, making it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms.
We then validate the model with a hardware implementation on real radio signals, detailed in Section~\ref{sec:4:gnuradio}.
%
We conclude this Chapter by presenting in Section~\ref{sec:4:retransmissions} an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.

@@

% ----------------------------------------------------------------------------
\section{Extension of our model to account for retransmissions}
\label{sec:4:retransmissions}
% ----------------------------------------------------------------------

@@

We give in this Section conclusions about the two models presented above and about our demonstration, as well as many directions of future works, some of which are currently being studied by researchers of our team.

@@

% ----------------------------------------------------------------------
\paragraph{Summary of the first model}
\label{sub:41:conclusion}

@@

% ----------------------------------------------------------------------
\paragraph{Conclusions taken from the demonstration}
\label{sub:42:conclusionFromDemonstration}

@@

% ----------------------------------------------------------------------
\paragraph{Conclusion about this second model}
\label{sub:43:conclusion}

@@

Future works related to this Chapter include the following directions.

\paragraph{Possible extensions of the first model}

The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices had different transmission pattern, and less easily with one probability per device. Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
In this whole Chapter~\ref{chapter:4}, we prefer to consider that all devices have the same activation probability, to keep the notation as simple as possible.

Moreover, for sake of simplicity we supposed that all devices use the same standard.
Future works could consider more realistic interference scenarios and IoT networks, with, \eg, non-slotted time, more than one base station etc.

@@

\paragraph{Extensions of the demonstration}

@@

\paragraph{Possible extensions for the second model}

The utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works. Among them, the non-stationarity of the channel occupancy caused by the learning policy employed by the IoT devices.
%
For that end, modifications of MAB algorithms have been proposed, such as Sliding-Window-\UCB{} or Discounted-\UCB{} \cite{Garivier11UCBDiscount}
or more recently M-\UCB{} \cite{CaoZhenKvetonXie18},
or more recently GLR-\UCB{} \cite{Besson2019GLRT} which is presented in Chapter~\ref{chapter:6},
that nevertheless have not been explored for the targeted problem.
Chapter~\ref{chapter:6} is focusing on this direction, but we did not have enough time to explore the possible applications of MAB policies designed for non-stationary problems to the model with retransmissions presented in this Chapter.

@@

\newpage
% ----------------------------------------------------------------------------
\section{Appendix}
\label{sec:4:appendix}
% ----------------------------------------------------------------------------

We include here a missing proof from Section~\ref{sec:4:firstModel}.

>>>>

More precisely, we start by motivating the idea of independent ``selfish'' learning for IoT devices in Section~\ref{sec:4:motivations}.
We then present in Section~\ref{sec:4:firstModel} a model where independent devices use an IoT network composed of different channels, occupied by static devices.
Devices do not use any sensing, but use acknowledgements sent by the Base Station to know if their uplink messages were successfully received.
We consider the case when many independent ``dynamic'' devices all have a small probability of communicating at every discrete time. If they use simple MAB algorithms in a decentralized and independent fashion (what we call ``selfish'' learning),
we show in simulations that in the case of a non-uniform repartition of the static devices in the channels, the dynamic devices can drastically improve the network efficiency.
%
On a simple example, we present in Section~\ref{sec:4:gnuradio} a demonstration that implements and validates the proposed approach, using real wireless radio hardware.

Then we present in Section~\ref{sec:4:retransmissions} a second model, extending our first model, in which devices can retransmit their messages for a fixed number of maximum retransmission, like in the ALOHA protocol.
In this second model, we compare five different heuristics, all based on the UCB algorithm, and numerical simulation show that all the non-naive heuristics succeed in significantly improving the network efficiency.
Interestingly, the simplest heuristic is found to be the most efficient one, as its simplicity implies a faster learning and convergence time.

@@

This chapter is mainly based on the following publications: \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.

@@

Efficient Medium Access (MAC) policies allow devices to avoid interfering traffic and can significantly reduce the spectrum contention problem in unlicensed bands.
As end-devices battery life is a key constraint of IoT networks,
this leads to IoT protocols using as low signaling overhead as possible and simple ALOHA-based mechanisms.
%
In this chapter, we analyze the performance of Multi-Armed Bandits (MAB) algorithms, used in combination with a time-frequency slotted ALOHA-based protocol.
We consider the Upper-Confidence Bound (\UCB) \cite{Auer02}, and the Thompson-Sampling (TS) algorithms \cite{Thompson33,AgrawalGoyal11,
Kaufmann12Thompson}, for the first model. For the demonstration as well as for the second model, we preferred to focus on heuristics based on the simplest algorithm (\ie, \UCB), to give a clear presentation of the different ideas explored for solve the problem of learning in order to retransmit efficiently.

@@

We present in Section~\ref{sec:4:firstModel} how the MAB algorithms can be used in a unlicensed but frequency- and time-slotted IoT network.
Several devices are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously and their activation processes are random, their behavior is not stationary.
As far as we know, we provide the first practical study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent devices in the network, which makes the environment ``highly'' not stationary, making it very hard to give mathematical proofs of  convergence and efficiency for bandit algorithms.
We then validate the model with a hardware implementation on real radio signals, detailed in Section~\ref{sec:4:gnuradio}.
%
We conclude this chapter by presenting in Section~\ref{sec:4:retransmissions} an extension of this model to take into account another aspect of the ALOHA protocol, that is the possibility for dynamic devices to retransmit their packet if the \emph{Ack} was not received.

@@

% ----------------------------------------------------------------------------
\section{Extending the model to account for retransmissions}
\label{sec:4:retransmissions}
% ----------------------------------------------------------------------

@@

We give in this section conclusions about the two models presented above and about our demonstration, as well as many directions of future works, some of which are currently being studied by researchers of our team.

\TODOL{Je devrais réécrire toute cette conclusion, si j'ai le temps, pour faire mieux qu'une suite de 3 conclusions et 3 future works, et être plus concis et moins ``short sighted'' comme disait Emilie.}

@@

% ----------------------------------------------------------------------
% \paragraph{Summary of the first model}
\label{sub:41:conclusion}

@@

% ----------------------------------------------------------------------
% \paragraph{Conclusions taken from the demonstration}
\label{sub:42:conclusionFromDemonstration}

@@

% ----------------------------------------------------------------------
% \paragraph{Conclusion about this second model}
\label{sub:43:conclusion}

@@

Future works related to this chapter include the following directions.

% \paragraph{Possible extensions of the first model}

The first model presented in Section~\ref{sec:4:firstModel} could easily be generalized with two probabilities $p_S$ and $p_D$, if static and dynamic devices have different transmission pattern, and less easily with one probability per device. Also, other emission pattern could be considered, instead of a Bernoulli process for each user.
In this whole Chapter~\ref{chapter:4}, we prefer to consider that all devices have the same activation probability, to keep the notation as simple as possible.
%
Moreover, for sake of simplicity we supposed that all devices use the same standard.
Future works could consider more realistic interference scenarios and IoT networks, with, \eg, non-slotted time, more than one base station etc.

@@

% \paragraph{Extensions of the demonstration}

@@

% \paragraph{Possible extensions for the second model}

Finally, the utility and impact of the proposed approaches for LPWA networks motivates us to address several subjects as future works. Among them, the non-stationarity of the channel occupancy caused by the learning policy employed by the IoT devices.
%
For that end, modifications of MAB algorithms have been proposed, such as Sliding-Window-\UCB{} or Discounted-\UCB{} \cite{Garivier11UCBDiscount}
or more recently M-\UCB{} \cite{CaoZhenKvetonXie18},
or more recently GLR-\UCB{} \cite{Besson2019GLRT} which is presented in Chapter~\ref{chapter:6},
that nevertheless have not been explored for the targeted problem.
Chapter~\ref{chapter:6} is focusing on this direction, but we did not have enough time to explore the possible applications of MAB policies designed for non-stationary problems to the model with retransmissions presented in this Chapter.

@@

\subsection*{Summary of this chapter}

To sum-up, we focussed in this chapter on models of IoT networks, and we proposed to use classical stationary multi-armed bandit learning algorithms implemented in a selfish and decentralized manner by each of the dynamic devices of the IoT network.
We presented two models of wireless IoT network without sensing, inspired by the ALOHA protocol, and with or without retransmission of uplink packages in case of collisions.
In both cases, we try to model the existing standards, like the LoRa standard, and we demonstrated the efficiency of the proposed MAB-based approach in both numerical simulations and empirical measurements on real wireless radio signals.
%
On one hand, a satisfying message is that this learning-based approach seems to be very efficient, as it allows the IoT devices to automatically and independently increase their successful transmission rate.
On the other hand, it is quite surprising that stochastic MAB algorithms can be of any use in such non-stationary applications.
Sadly, analyzing our model with thousands of independent devices communicating and learning in their own (random) time scale turned out to be of extreme difficulty.

\paragraph{Multi-Player MAB with \emph{collision avoidance}}

We are actually able to analyze a simpler model, if we assume to have at most $K$ devices with a transmission probability of $p=1$.
Instead of the direction pursued in this chapter, another possibility is to consider a \emph{multi-player MAB} model to describe our problem, as proposed by \cite{Zhao10}.
%
The main difference between the two models is the fact that in this chapter, devices do not transmit their messages at every time step, but following a random activation process (with a fixed transmission probability $p < 1$).
If static and dynamic devices that have to transmit at a fixed time are denoted \emph{active devices},
and their random activation patter makes the number of active devices a random variable.
Analyzing multi-player MAB models under this hypothesis is much harder, and is left as a future work.
%
We study this other direction in the next Chapter~\ref{chapter:5}.

% In that case, the static and dynamic devices effect is decoupled, and arms only model the availability of the channels in the absence of dynamic devices: they are \iid{} with mean $\mu_i = 1 - p S_i$.
% Moreover, dynamic devices are usually assumed to be able to \emph{sense} a channel before sending \cite{Zhao10}, and so communicate only if no static device is detected on the channel.
% The smart devices try to learn the arms with highest means, while coordinating to choose different arms, \ie, avoid collisions in their choice, in a decentralized manner.
% However, in this model it is assumed that the multiple agents can know that they experienced a collision with another agent, which is non-realistic for our problem at stake, as our model of smart device cannot do sensing nor differentiate collisions between smart and non-smart devices.

\newpage
% ----------------------------------------------------------------------------
\section{Appendix}
\label{sec:4:appendix}
% \section{Proof of Proposition \ref{prop:41:Lagrangian}}
\label{sec:5:proofLagrangian}
% ----------------------------------------------------------------------------

% We include here a missing proof from Section~\ref{sec:4:firstModel}.
